---
title: "ANCOVA in Vocabulario Ensinado (Vocabulario Ensinado)"
author: Geiser C. Challco <geiser@alumni.usp.br>
comment: Teste ANCOVA para determinar se houve diferenças significativas
         na Vocabulario Ensinado (medido usando pre- e pos-testes).
         
         ANCOVA test to determine whether there were significant differences
         in Vocabulario Ensinado (measured using pre- and post-tests).
         
         Author - Geiser C. Challco <geiser@alumni.usp.br>
         
         Shiny-Statistic is distributed in the hope that it will be useful,
         but WITHOUT ANY WARRANTY; without even the implied warranty of
         MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
         GNU General Public License for more details.
         
         You should have received a copy of the GNU General Public License.
         This file is  generate using Shiny-Statistic app (https://github.com/geiser/rshinystatistics)
         If not, see <https://www.gnu.org/licenses/>.
output:
  github_document:
    toc: true
  word_document:
    toc: true
  html_document:
    toc: true
fontsize: 10pt
---


**NOTE**:

 - Teste ANCOVA para determinar se houve diferenças significativas
no Vocabulario Ensinado (medido usando pre- e pos-testes).
 - ANCOVA test to determine whether there were significant differences
in Vocabulario Ensinado (measured using pre- and post-tests).


```{r setup, include=FALSE}
# Install and Load Packages
if (!'remotes' %in% rownames(installed.packages())) install.packages('remotes')
if (!"rshinystatistics" %in% rownames(installed.packages())) {
  remotes::install_github("geiser/rshinystatistics")
} else if (packageVersion("rshinystatistics") < "0.0.0.9900") {
  remotes::install_github("geiser/rshinystatistics")
}

wants <- c('ggplot2','ggpubr','rshinystatistics','utils','randomcoloR')
has <- wants %in% rownames(installed.packages())
if (any(!has)) install.packages(wants[!has])

library(readxl)

library(shiny)
library(esquisse)
library(scales)
library(knitr)
library(rmarkdown)

library(utils)
library(ggpubr)
library(ggplot2)
library(randomcoloR)

library(dplyr)

library(rstatix)
library(rshinystatistics)
```

```{r, include=FALSE, purl=FALSE}
options(knitr.kable.NA = '')
```

```{r, include=FALSE}
# Setting Initial Variables
dv = "vocab.teach"
dv.pos = "vocab.teach.pos"
dv.pre = "vocab.teach.pre"

fatores2 <- c("Sexo","Zona","Cor.Raca","Serie","vocab.teach.quintile")
lfatores2 <- as.list(fatores2)
names(lfatores2) <- fatores2

fatores1 <- c("grupo", fatores2)
lfatores1 <- as.list(fatores1)
names(lfatores1) <- fatores1

lfatores <- c(lfatores1)

# ..

gdat <- read_excel("../data/data.xlsx", sheet = "sumary")
gdat <- gdat[which(is.na(gdat$Necessidade.Deficiencia) & !is.na(gdat$WG.Grupo)),]



dat <- gdat
dat$grupo <- factor(dat[["WG.Grupo"]], c("Controle","Experimental"))
for (coln in c(names(lfatores))) {
  dat[[coln]] <- factor(dat[[coln]], sort(unique(dat[[coln]])[!is.na(unique(dat[[coln]]))]))
}
dat <- dat[which(!is.na(dat[[dv.pre]]) & !is.na(dat[[dv.pos]])),]
dat <- dat[,c("id",names(lfatores),dv.pre,dv.pos)]

dat.long <- rbind(dat, dat)
dat.long$time <- c(rep("pre", nrow(dat)), rep("pos", nrow(dat)))
dat.long$time <- factor(dat.long$time, c("pre","pos"))
dat.long[[dv]] <- c(dat[[dv.pre]], dat[[dv.pos]])

color <- list()
color[["prepost"]] = c("#ffee65","#f28e2B")
color[["grupo"]] = c("#bcbd22","#008000")
color[["Sexo"]] = c("#FF007F","#4D4DFF")
color[["Zona"]] = c("#AA00FF","#00CCCC")
color[["Cor.Raca"]] = c(
  "Parda"="#b97100","Indígena"="#9F262F",
  "Branca"="#87c498", "Preta"="#848283","Amarela"="#D6B91C"
)

color[["grupo:Sexo"]] = c(
  "Controle:F"="#ff99cb", "Controle:M"="#b7b7ff",
  "Experimental:F"="#FF007F", "Experimental:M"="#4D4DFF",
  "Controle.F"="#ff99cb", "Controle.M"="#b7b7ff",
  "Experimental.F"="#FF007F", "Experimental.M"="#4D4DFF"
)
color[["grupo:Zona"]] = c(
  "Controle:Rural"="#b2efef","Controle:Urbana"="#e5b2ff",
  "Experimental:Rural"="#00CCCC", "Experimental:Urbana"="#AA00FF",
  "Controle.Rural"="#b2efef","Controle.Urbana"="#e5b2ff",
  "Experimental.Rural"="#00CCCC", "Experimental.Urbana"="#AA00FF"
)
color[["grupo:Cor.Raca"]] = c(
    "Controle:Parda"="#e3c699", "Experimental:Parda"="#b97100",
    "Controle:Indígena"="#e2bdc0", "Experimental:Indígena"="#9F262F",
    "Controle:Branca"="#c0e8cb", "Experimental:Branca"="#87c498",
    "Controle:Preta"="#dad9d9", "Experimental:Preta"="#848283",
    "Controle:Amarela"="#eee3a4", "Experimental:Amarela"="#D6B91C",
    
    "Controle.Parda"="#e3c699", "Experimental.Parda"="#b97100",
    "Controle.Indígena"="#e2bdc0", "Experimental.Indígena"="#9F262F",
    "Controle.Branca"="#c0e8cb", "Experimental.Branca"="#87c498",
    "Controle.Preta"="#dad9d9", "Experimental.Preta"="#848283",
    "Controle.Amarela"="#eee3a4", "Experimental.Amarela"="#D6B91C"
)


for (coln in c("vocab","score.tde",
               "TFL.lidas.per.min","TFL.corretas.per.min","leitura.compreensao")) {
  color[[paste0(coln,".quintile")]] = c("#BF0040","#FF0000","#800080","#0000FF","#4000BF")
  color[[paste0("grupo:",coln,".quintile")]] = c(
    "Experimental.1st quintile"="#BF0040", "Controle.1st quintile"="#d8668c",
    "Experimental.2nd quintile"="#FF0000", "Controle.2nd quintile"="#ff7f7f",
    "Experimental.3rd quintile"="#8fce00", "Controle.3rd quintile"="#ddf0b2",
    "Experimental.4th quintile"="#0000FF", "Controle.4th quintile"="#b2b2ff",
    "Experimental.5th quintile"="#4000BF", "Controle.5th quintile"="#b299e5",
    
    "Experimental:1st quintile"="#BF0040", "Controle:1st quintile"="#d8668c",
    "Experimental:2nd quintile"="#FF0000", "Controle:2nd quintile"="#ff7f7f",
    "Experimental:3rd quintile"="#8fce00", "Controle:3rd quintile"="#ddf0b2",
    "Experimental:4th quintile"="#0000FF", "Controle:4th quintile"="#b2b2ff",
    "Experimental:5th quintile"="#4000BF", "Controle:5th quintile"="#b299e5")
}


for (f in c("grupo", names(lfatores))) {
  if (is.null(color[[f]]) && length(unique(dat[[f]])) > 0) 
      color[[f]] <- distinctColorPalette(length(unique(dat[[f]])))
}
for (f in c(fatores2)) {
  if (is.null(color[[paste0("grupo:",f)]]) && length(unique(dat[[f]])) > 0)
    color[[paste0("grupo:",f)]] <- distinctColorPalette(length(unique(dat[["grupo"]]))*length(unique(dat[[f]])))
}

ldat <- list()
laov <- list()
lpwc <- list()
lemms <- list()
```

# Descriptive Statistics of Initial Data

```{r}
df <- get.descriptives(dat, c(dv.pre, dv.pos), c("grupo"), 
                       include.global = T, symmetry.test = T, normality.test = F)
df <- plyr::rbind.fill(
  df, do.call(plyr::rbind.fill, lapply(lfatores2, FUN = function(f) {
    if (nrow(dat) > 0 && sum(!is.na(unique(dat[[f]]))) > 1)
      get.descriptives(dat, c(dv.pre,dv.pos), c("grupo", f),
                       symmetry.test = T, normality.test = F)
    }))
)
(df <- df[,c(fatores1[fatores1 %in% colnames(df)],"variable",
             colnames(df)[!colnames(df) %in% c(fatores1,"variable")])])
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```




# ANCOVA and Pairwise for one factor: **grupo**

## Without remove non-normal data

```{r}
pdat = remove_group_data(dat[!is.na(dat[["grupo"]]),], "vocab.teach.pos", "grupo")

pdat.long <- rbind(pdat[,c("id","grupo")], pdat[,c("id","grupo")])
pdat.long[["time"]] <- c(rep("pre", nrow(pdat)), rep("pos", nrow(pdat)))
pdat.long[["time"]] <- factor(pdat.long[["time"]], c("pre","pos"))
pdat.long[["vocab.teach"]] <- c(pdat[["vocab.teach.pre"]], pdat[["vocab.teach.pos"]])

aov = anova_test(pdat, vocab.teach.pos ~ vocab.teach.pre + grupo)
laov[["grupo"]] <- get_anova_table(aov)
```


```{r}
pwc <- emmeans_test(pdat, vocab.teach.pos ~ grupo, covariate = vocab.teach.pre,
                    p.adjust.method = "bonferroni")
```

```{r}
pwc.long <- emmeans_test(dplyr::group_by_at(pdat.long, "grupo"),
                          vocab.teach ~ time,
                          p.adjust.method = "bonferroni")
lpwc[["grupo"]] <- plyr::rbind.fill(pwc, pwc.long)
```

```{r}
ds <- get.descriptives(pdat, "vocab.teach.pos", "grupo", covar = "vocab.teach.pre")
ds <- merge(ds[ds$variable != "vocab.teach.pre",],
            ds[ds$variable == "vocab.teach.pre", !colnames(ds) %in% c("variable")],
            by = "grupo", all.x = T, suffixes = c("", ".vocab.teach.pre"))
ds <- merge(get_emmeans(pwc), ds, by = "grupo", suffixes = c(".emms", ""))
ds <- ds[,c("grupo","n","mean.vocab.teach.pre","se.vocab.teach.pre","mean","se","emmean","se.emms")]

colnames(ds) <- c("grupo", "N", paste0(c("M","SE")," (pre)"),
                  paste0(c("M","SE"), " (unadj)"), paste0(c("M", "SE"), " (adj)"))

lemms[["grupo"]] <- ds
```

## Computing ANCOVA and PairWise After removing non-normal data (OK)

```{r}
wdat = pdat 

res = residuals(lm(vocab.teach.pos ~ vocab.teach.pre + grupo, data = wdat))
non.normal = getNonNormal(res, wdat$id, plimit = 0.05)

wdat = wdat[!wdat$id %in% non.normal,]

wdat.long <- rbind(wdat[,c("id","grupo")], wdat[,c("id","grupo")])
wdat.long[["time"]] <- c(rep("pre", nrow(wdat)), rep("pos", nrow(wdat)))
wdat.long[["time"]] <- factor(wdat.long[["time"]], c("pre","pos"))
wdat.long[["vocab.teach"]] <- c(wdat[["vocab.teach.pre"]], wdat[["vocab.teach.pos"]])

ldat[["grupo"]] = wdat

(non.normal)
```

```{r}
aov = anova_test(wdat, vocab.teach.pos ~ vocab.teach.pre + grupo)
laov[["grupo"]] <- merge(get_anova_table(aov), laov[["grupo"]], by="Effect", suffixes = c("","'"))

(df = get_anova_table(aov))
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

```{r, echo=FALSE}
pwc <- emmeans_test(wdat, vocab.teach.pos ~ grupo, covariate = vocab.teach.pre,
                    p.adjust.method = "bonferroni")
```

```{r, echo=FALSE, purl=FALSE}
kable(pwc, digits = 3)
```

```{r}
pwc.long <- emmeans_test(dplyr::group_by_at(wdat.long, "grupo"),
                         vocab.teach ~ time,
                         p.adjust.method = "bonferroni")
lpwc[["grupo"]] <- merge(plyr::rbind.fill(pwc, pwc.long), lpwc[["grupo"]], by=c("grupo","term",".y.","group1","group2"), suffixes = c("","'"))
```

```{r, echo=FALSE, purl=FALSE}
kable(pwc.long, digits = 3)
```


```{r}
ds <- get.descriptives(wdat, "vocab.teach.pos", "grupo", covar = "vocab.teach.pre")
ds <- merge(ds[ds$variable != "vocab.teach.pre",],
            ds[ds$variable == "vocab.teach.pre", !colnames(ds) %in% c("variable")],
            by = "grupo", all.x = T, suffixes = c("", ".vocab.teach.pre"))
ds <- merge(get_emmeans(pwc), ds, by = "grupo", suffixes = c(".emms", ""))
ds <- ds[,c("grupo","n","mean.vocab.teach.pre","se.vocab.teach.pre","mean","se","emmean","se.emms")]

colnames(ds) <- c("grupo", "N", paste0(c("M","SE")," (pre)"),
                  paste0(c("M","SE"), " (unadj)"), paste0(c("M", "SE"), " (adj)"))

lemms[["grupo"]] <- merge(ds, lemms[["grupo"]], by=c("grupo"), suffixes = c("","'"))
```

```{r, echo=FALSE, purl=FALSE}
kable(ds, digits = 3)
```


### Plots for ancova

```{r}
plots <- oneWayAncovaPlots(
  wdat, "vocab.teach.pos", "grupo", aov, list("grupo"=pwc), addParam = c("mean_se"),
  font.label.size=10, step.increase=0.05, p.label="p.adj",
  subtitle = which(aov$Effect == "grupo"))
```

```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(nrow(plots[["grupo"]]$data)))
  plots[["grupo"]] + ggplot2::scale_color_manual(values = color[["grupo"]])
```


```{r}
plots <- oneWayAncovaBoxPlots(
  wdat, "vocab.teach.pos", "grupo", aov, pwc, covar = "vocab.teach.pre",
  theme = "classic", color = color[["grupo"]],
  subtitle = which(aov$Effect == "grupo"))
```

```{r, dpi=300, fig.width=8, fig.height=6}
if (length(unique(wdat[["grupo"]])) > 1)
  plots[["grupo"]] + ggplot2::ylab("Vocabulario Ensinado") + ggplot2::scale_x_discrete(labels=c('pre', 'pos'))
```


```{r}
if (length(unique(wdat.long[["grupo"]])) > 1)
  plots <- oneWayAncovaBoxPlots(
    wdat.long, "vocab.teach", "grupo", aov, pwc.long, pre.post = "time",
    theme = "classic", color = color$prepost)
```

```{r, dpi=300, fig.width=10, fig.height=6}
if (length(unique(wdat.long[["grupo"]])) > 1)
  plots[["grupo"]] + ggplot2::ylab("Vocabulario Ensinado")
```


### Checking linearity assumption

```{r, dpi=300, fig.width=12, fig.height=8}
ggscatter(wdat, x = "vocab.teach.pre", y = "vocab.teach.pos", size = 0.5,
          color = "grupo", add = "reg.line")+
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~"), color = grupo)
  )
```

### Checking normality and homogeneity

```{r}
res <- augment(lm(vocab.teach.pos ~ vocab.teach.pre + grupo, data = wdat))
```

```{r}
shapiro_test(res$.resid)
```

```{r}
levene_test(res, .resid ~ grupo)
```


# ANCOVA and Pairwise for two factors **grupo:Sexo**

## Without remove non-normal data

```{r}
pdat = remove_group_data(dat[!is.na(dat[["grupo"]]) & !is.na(dat[["Sexo"]]),], "vocab.teach.pos", c("grupo","Sexo"))

pdat.long <- rbind(pdat[,c("id","grupo","Sexo")], pdat[,c("id","grupo","Sexo")])
pdat.long[["time"]] <- c(rep("pre", nrow(pdat)), rep("pos", nrow(pdat)))
pdat.long[["time"]] <- factor(pdat.long[["time"]], c("pre","pos"))
pdat.long[["vocab.teach"]] <- c(pdat[["vocab.teach.pre"]], pdat[["vocab.teach.pos"]])

aov = anova_test(pdat, vocab.teach.pos ~ vocab.teach.pre + grupo*Sexo)
laov[["grupo:Sexo"]] <- get_anova_table(aov)
```

```{r}
pwcs <- list()
pwcs[["Sexo"]] <- emmeans_test(
  group_by(pdat, grupo), vocab.teach.pos ~ Sexo,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")
pwcs[["grupo"]] <- emmeans_test(
  group_by(pdat, Sexo), vocab.teach.pos ~ grupo,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")

pwc <- plyr::rbind.fill(pwcs[["grupo"]], pwcs[["Sexo"]])
pwc <- pwc[,c("grupo","Sexo", colnames(pwc)[!colnames(pwc) %in% c("grupo","Sexo")])]
```

```{r}
pwc.long <- emmeans_test(dplyr::group_by_at(pdat.long, c("grupo","Sexo")),
                         vocab.teach ~ time,
                         p.adjust.method = "bonferroni")
lpwc[["grupo:Sexo"]] <- plyr::rbind.fill(pwc, pwc.long)
```


```{r}
ds <- get.descriptives(pdat, "vocab.teach.pos", c("grupo","Sexo"), covar = "vocab.teach.pre")
ds <- merge(ds[ds$variable != "vocab.teach.pre",],
            ds[ds$variable == "vocab.teach.pre", !colnames(ds) %in% c("variable")],
            by = c("grupo","Sexo"), all.x = T, suffixes = c("", ".vocab.teach.pre"))
ds <- merge(get_emmeans(pwcs[["grupo"]]), ds, by = c("grupo","Sexo"), suffixes = c(".emms", ""))
ds <- ds[,c("grupo","Sexo","n","mean.vocab.teach.pre","se.vocab.teach.pre","mean","se","emmean","se.emms")]

colnames(ds) <- c("grupo","Sexo", "N", paste0(c("M","SE")," (pre)"),
                  paste0(c("M","SE"), " (unadj)"), paste0(c("M", "SE"), " (adj)"))

lemms[["grupo:Sexo"]] <- ds
```

## Computing ANCOVA and PairWise After removing non-normal data (OK)

```{r}
wdat = pdat 

res = residuals(lm(vocab.teach.pos ~ vocab.teach.pre + grupo*Sexo, data = wdat))
non.normal = getNonNormal(res, wdat$id, plimit = 0.05)

wdat = wdat[!wdat$id %in% non.normal,]

wdat.long <- rbind(wdat[,c("id","grupo","Sexo")], wdat[,c("id","grupo","Sexo")])
wdat.long[["time"]] <- c(rep("pre", nrow(wdat)), rep("pos", nrow(wdat)))
wdat.long[["time"]] <- factor(wdat.long[["time"]], c("pre","pos"))
wdat.long[["vocab.teach"]] <- c(wdat[["vocab.teach.pre"]], wdat[["vocab.teach.pos"]])


ldat[["grupo:Sexo"]] = wdat

(non.normal)
```

```{r}
aov = anova_test(wdat, vocab.teach.pos ~ vocab.teach.pre + grupo*Sexo)
laov[["grupo:Sexo"]] <- merge(get_anova_table(aov), laov[["grupo:Sexo"]], by="Effect", suffixes = c("","'"))

(df = get_anova_table(aov))
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

```{r}
pwcs <- list()
pwcs[["Sexo"]] <- emmeans_test(
  group_by(wdat, grupo), vocab.teach.pos ~ Sexo,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")
pwcs[["grupo"]] <- emmeans_test(
  group_by(wdat, Sexo), vocab.teach.pos ~ grupo,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")

pwc <- plyr::rbind.fill(pwcs[["grupo"]], pwcs[["Sexo"]])
pwc <- pwc[,c("grupo","Sexo", colnames(pwc)[!colnames(pwc) %in% c("grupo","Sexo")])]
```

```{r, echo=FALSE, purl=FALSE}
kable(pwc, digits = 3)
```

```{r}
pwc.long <- emmeans_test(dplyr::group_by_at(wdat.long, c("grupo","Sexo")),
                         vocab.teach ~ time,
                         p.adjust.method = "bonferroni")
lpwc[["grupo:Sexo"]] <- merge(plyr::rbind.fill(pwc, pwc.long), lpwc[["grupo:Sexo"]], by=c("grupo","Sexo","term",".y.","group1","group2"), suffixes = c("","'"))
```

```{r, echo=FALSE, purl=FALSE}
kable(pwc.long, digits = 3)
```

```{r}
ds <- get.descriptives(wdat, "vocab.teach.pos", c("grupo","Sexo"), covar = "vocab.teach.pre")
ds <- merge(ds[ds$variable != "vocab.teach.pre",],
            ds[ds$variable == "vocab.teach.pre", !colnames(ds) %in% c("variable")],
            by = c("grupo","Sexo"), all.x = T, suffixes = c("", ".vocab.teach.pre"))
ds <- merge(get_emmeans(pwcs[["grupo"]]), ds, by = c("grupo","Sexo"), suffixes = c(".emms", ""))
ds <- ds[,c("grupo","Sexo","n","mean.vocab.teach.pre","se.vocab.teach.pre","mean","se","emmean","se.emms")]

colnames(ds) <- c("grupo","Sexo", "N", paste0(c("M","SE")," (pre)"),
                  paste0(c("M","SE"), " (unadj)"), paste0(c("M", "SE"), " (adj)"))

lemms[["grupo:Sexo"]] <- merge(ds, lemms[["grupo:Sexo"]], by=c("grupo","Sexo"), suffixes = c("","'"))
```

```{r, echo=FALSE, purl=FALSE}
kable(ds, digits = 3)
```


### Plots for ancova

```{r}
plots <- twoWayAncovaPlots(
  wdat, "vocab.teach.pos", c("grupo","Sexo"), aov, pwcs, addParam = c("mean_se"),
  font.label.size=10, step.increase=0.05, p.label="p.adj",
  subtitle = which(aov$Effect == "grupo:Sexo"))
```

```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(plots[["grupo"]]))
  plots[["grupo"]] + ggplot2::scale_color_manual(values = color[["Sexo"]])
```


```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(plots[["Sexo"]]))
  plots[["Sexo"]] + ggplot2::scale_color_manual(values = color[["grupo"]])
```


```{r}
plots <- twoWayAncovaBoxPlots(
  wdat, "vocab.teach.pos", c("grupo","Sexo"), aov, pwcs, covar = "vocab.teach.pre",
  theme = "classic", color = color[["grupo:Sexo"]],
  subtitle = which(aov$Effect == "grupo:Sexo"))
```

```{r, dpi=300, fig.width=8, fig.height=6}
plots[["grupo:Sexo"]] + ggplot2::ylab("Vocabulario Ensinado") + ggplot2::scale_x_discrete(labels=c('pre', 'pos'))
```


```{r}
plots <- twoWayAncovaBoxPlots(
  wdat.long, "vocab.teach", c("grupo","Sexo"), aov, pwc.long, pre.post = "time",
  theme = "classic", color = color$prepost)
```

```{r, dpi=300, fig.width=14, fig.height=6}
plots[["grupo:Sexo"]] + ggplot2::ylab("Vocabulario Ensinado")
```


### Checking linearity assumption

```{r, dpi=300, fig.width=12, fig.height=8}
ggscatter(wdat, x = "vocab.teach.pre", y = "vocab.teach.pos", size = 0.5,
          facet.by = c("grupo","Sexo"), add = "reg.line")+
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~"))
  )
```

### Checking normality and homogeneity

```{r}
res <- augment(lm(vocab.teach.pos ~ vocab.teach.pre + grupo*Sexo, data = wdat))
```

```{r}
shapiro_test(res$.resid)
```

```{r}
levene_test(res, .resid ~ grupo*Sexo)
```


# ANCOVA and Pairwise for two factors **grupo:Zona**

## Without remove non-normal data

```{r}
pdat = remove_group_data(dat[!is.na(dat[["grupo"]]) & !is.na(dat[["Zona"]]),], "vocab.teach.pos", c("grupo","Zona"))

pdat.long <- rbind(pdat[,c("id","grupo","Zona")], pdat[,c("id","grupo","Zona")])
pdat.long[["time"]] <- c(rep("pre", nrow(pdat)), rep("pos", nrow(pdat)))
pdat.long[["time"]] <- factor(pdat.long[["time"]], c("pre","pos"))
pdat.long[["vocab.teach"]] <- c(pdat[["vocab.teach.pre"]], pdat[["vocab.teach.pos"]])

aov = anova_test(pdat, vocab.teach.pos ~ vocab.teach.pre + grupo*Zona)
laov[["grupo:Zona"]] <- get_anova_table(aov)
```

```{r}
pwcs <- list()
pwcs[["Zona"]] <- emmeans_test(
  group_by(pdat, grupo), vocab.teach.pos ~ Zona,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")
pwcs[["grupo"]] <- emmeans_test(
  group_by(pdat, Zona), vocab.teach.pos ~ grupo,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")

pwc <- plyr::rbind.fill(pwcs[["grupo"]], pwcs[["Zona"]])
pwc <- pwc[,c("grupo","Zona", colnames(pwc)[!colnames(pwc) %in% c("grupo","Zona")])]
```

```{r}
pwc.long <- emmeans_test(dplyr::group_by_at(pdat.long, c("grupo","Zona")),
                         vocab.teach ~ time,
                         p.adjust.method = "bonferroni")
lpwc[["grupo:Zona"]] <- plyr::rbind.fill(pwc, pwc.long)
```


```{r}
ds <- get.descriptives(pdat, "vocab.teach.pos", c("grupo","Zona"), covar = "vocab.teach.pre")
ds <- merge(ds[ds$variable != "vocab.teach.pre",],
            ds[ds$variable == "vocab.teach.pre", !colnames(ds) %in% c("variable")],
            by = c("grupo","Zona"), all.x = T, suffixes = c("", ".vocab.teach.pre"))
ds <- merge(get_emmeans(pwcs[["grupo"]]), ds, by = c("grupo","Zona"), suffixes = c(".emms", ""))
ds <- ds[,c("grupo","Zona","n","mean.vocab.teach.pre","se.vocab.teach.pre","mean","se","emmean","se.emms")]

colnames(ds) <- c("grupo","Zona", "N", paste0(c("M","SE")," (pre)"),
                  paste0(c("M","SE"), " (unadj)"), paste0(c("M", "SE"), " (adj)"))

lemms[["grupo:Zona"]] <- ds
```

## Computing ANCOVA and PairWise After removing non-normal data (OK)

```{r}
wdat = pdat 

res = residuals(lm(vocab.teach.pos ~ vocab.teach.pre + grupo*Zona, data = wdat))
non.normal = getNonNormal(res, wdat$id, plimit = 0.05)

wdat = wdat[!wdat$id %in% non.normal,]

wdat.long <- rbind(wdat[,c("id","grupo","Zona")], wdat[,c("id","grupo","Zona")])
wdat.long[["time"]] <- c(rep("pre", nrow(wdat)), rep("pos", nrow(wdat)))
wdat.long[["time"]] <- factor(wdat.long[["time"]], c("pre","pos"))
wdat.long[["vocab.teach"]] <- c(wdat[["vocab.teach.pre"]], wdat[["vocab.teach.pos"]])


ldat[["grupo:Zona"]] = wdat

(non.normal)
```

```{r}
aov = anova_test(wdat, vocab.teach.pos ~ vocab.teach.pre + grupo*Zona)
laov[["grupo:Zona"]] <- merge(get_anova_table(aov), laov[["grupo:Zona"]], by="Effect", suffixes = c("","'"))

(df = get_anova_table(aov))
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

```{r}
pwcs <- list()
pwcs[["Zona"]] <- emmeans_test(
  group_by(wdat, grupo), vocab.teach.pos ~ Zona,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")
pwcs[["grupo"]] <- emmeans_test(
  group_by(wdat, Zona), vocab.teach.pos ~ grupo,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")

pwc <- plyr::rbind.fill(pwcs[["grupo"]], pwcs[["Zona"]])
pwc <- pwc[,c("grupo","Zona", colnames(pwc)[!colnames(pwc) %in% c("grupo","Zona")])]
```

```{r, echo=FALSE, purl=FALSE}
kable(pwc, digits = 3)
```

```{r}
pwc.long <- emmeans_test(dplyr::group_by_at(wdat.long, c("grupo","Zona")),
                         vocab.teach ~ time,
                         p.adjust.method = "bonferroni")
lpwc[["grupo:Zona"]] <- merge(plyr::rbind.fill(pwc, pwc.long), lpwc[["grupo:Zona"]], by=c("grupo","Zona","term",".y.","group1","group2"), suffixes = c("","'"))
```

```{r, echo=FALSE, purl=FALSE}
kable(pwc.long, digits = 3)
```

```{r}
ds <- get.descriptives(wdat, "vocab.teach.pos", c("grupo","Zona"), covar = "vocab.teach.pre")
ds <- merge(ds[ds$variable != "vocab.teach.pre",],
            ds[ds$variable == "vocab.teach.pre", !colnames(ds) %in% c("variable")],
            by = c("grupo","Zona"), all.x = T, suffixes = c("", ".vocab.teach.pre"))
ds <- merge(get_emmeans(pwcs[["grupo"]]), ds, by = c("grupo","Zona"), suffixes = c(".emms", ""))
ds <- ds[,c("grupo","Zona","n","mean.vocab.teach.pre","se.vocab.teach.pre","mean","se","emmean","se.emms")]

colnames(ds) <- c("grupo","Zona", "N", paste0(c("M","SE")," (pre)"),
                  paste0(c("M","SE"), " (unadj)"), paste0(c("M", "SE"), " (adj)"))

lemms[["grupo:Zona"]] <- merge(ds, lemms[["grupo:Zona"]], by=c("grupo","Zona"), suffixes = c("","'"))
```

```{r, echo=FALSE, purl=FALSE}
kable(ds, digits = 3)
```


### Plots for ancova

```{r}
plots <- twoWayAncovaPlots(
  wdat, "vocab.teach.pos", c("grupo","Zona"), aov, pwcs, addParam = c("mean_se"),
  font.label.size=10, step.increase=0.05, p.label="p.adj",
  subtitle = which(aov$Effect == "grupo:Zona"))
```

```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(plots[["grupo"]]))
  plots[["grupo"]] + ggplot2::scale_color_manual(values = color[["Zona"]])
```


```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(plots[["Zona"]]))
  plots[["Zona"]] + ggplot2::scale_color_manual(values = color[["grupo"]])
```


```{r}
plots <- twoWayAncovaBoxPlots(
  wdat, "vocab.teach.pos", c("grupo","Zona"), aov, pwcs, covar = "vocab.teach.pre",
  theme = "classic", color = color[["grupo:Zona"]],
  subtitle = which(aov$Effect == "grupo:Zona"))
```

```{r, dpi=300, fig.width=8, fig.height=6}
plots[["grupo:Zona"]] + ggplot2::ylab("Vocabulario Ensinado") + ggplot2::scale_x_discrete(labels=c('pre', 'pos'))
```


```{r}
plots <- twoWayAncovaBoxPlots(
  wdat.long, "vocab.teach", c("grupo","Zona"), aov, pwc.long, pre.post = "time",
  theme = "classic", color = color$prepost)
```

```{r, dpi=300, fig.width=14, fig.height=6}
plots[["grupo:Zona"]] + ggplot2::ylab("Vocabulario Ensinado")
```


### Checking linearity assumption

```{r, dpi=300, fig.width=12, fig.height=8}
ggscatter(wdat, x = "vocab.teach.pre", y = "vocab.teach.pos", size = 0.5,
          facet.by = c("grupo","Zona"), add = "reg.line")+
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~"))
  )
```

### Checking normality and homogeneity

```{r}
res <- augment(lm(vocab.teach.pos ~ vocab.teach.pre + grupo*Zona, data = wdat))
```

```{r}
shapiro_test(res$.resid)
```

```{r}
levene_test(res, .resid ~ grupo*Zona)
```


# ANCOVA and Pairwise for two factors **grupo:Cor.Raca**

## Without remove non-normal data

```{r}
pdat = remove_group_data(dat[!is.na(dat[["grupo"]]) & !is.na(dat[["Cor.Raca"]]),], "vocab.teach.pos", c("grupo","Cor.Raca"))

pdat.long <- rbind(pdat[,c("id","grupo","Cor.Raca")], pdat[,c("id","grupo","Cor.Raca")])
pdat.long[["time"]] <- c(rep("pre", nrow(pdat)), rep("pos", nrow(pdat)))
pdat.long[["time"]] <- factor(pdat.long[["time"]], c("pre","pos"))
pdat.long[["vocab.teach"]] <- c(pdat[["vocab.teach.pre"]], pdat[["vocab.teach.pos"]])

aov = anova_test(pdat, vocab.teach.pos ~ vocab.teach.pre + grupo*Cor.Raca)
laov[["grupo:Cor.Raca"]] <- get_anova_table(aov)
```

```{r}
pwcs <- list()
pwcs[["Cor.Raca"]] <- emmeans_test(
  group_by(pdat, grupo), vocab.teach.pos ~ Cor.Raca,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")
pwcs[["grupo"]] <- emmeans_test(
  group_by(pdat, Cor.Raca), vocab.teach.pos ~ grupo,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")

pwc <- plyr::rbind.fill(pwcs[["grupo"]], pwcs[["Cor.Raca"]])
pwc <- pwc[,c("grupo","Cor.Raca", colnames(pwc)[!colnames(pwc) %in% c("grupo","Cor.Raca")])]
```

```{r}
pwc.long <- emmeans_test(dplyr::group_by_at(pdat.long, c("grupo","Cor.Raca")),
                         vocab.teach ~ time,
                         p.adjust.method = "bonferroni")
lpwc[["grupo:Cor.Raca"]] <- plyr::rbind.fill(pwc, pwc.long)
```


```{r}
ds <- get.descriptives(pdat, "vocab.teach.pos", c("grupo","Cor.Raca"), covar = "vocab.teach.pre")
ds <- merge(ds[ds$variable != "vocab.teach.pre",],
            ds[ds$variable == "vocab.teach.pre", !colnames(ds) %in% c("variable")],
            by = c("grupo","Cor.Raca"), all.x = T, suffixes = c("", ".vocab.teach.pre"))
ds <- merge(get_emmeans(pwcs[["grupo"]]), ds, by = c("grupo","Cor.Raca"), suffixes = c(".emms", ""))
ds <- ds[,c("grupo","Cor.Raca","n","mean.vocab.teach.pre","se.vocab.teach.pre","mean","se","emmean","se.emms")]

colnames(ds) <- c("grupo","Cor.Raca", "N", paste0(c("M","SE")," (pre)"),
                  paste0(c("M","SE"), " (unadj)"), paste0(c("M", "SE"), " (adj)"))

lemms[["grupo:Cor.Raca"]] <- ds
```

## Computing ANCOVA and PairWise After removing non-normal data (OK)

```{r}
wdat = pdat 

res = residuals(lm(vocab.teach.pos ~ vocab.teach.pre + grupo*Cor.Raca, data = wdat))
non.normal = getNonNormal(res, wdat$id, plimit = 0.05)

wdat = wdat[!wdat$id %in% non.normal,]

wdat.long <- rbind(wdat[,c("id","grupo","Cor.Raca")], wdat[,c("id","grupo","Cor.Raca")])
wdat.long[["time"]] <- c(rep("pre", nrow(wdat)), rep("pos", nrow(wdat)))
wdat.long[["time"]] <- factor(wdat.long[["time"]], c("pre","pos"))
wdat.long[["vocab.teach"]] <- c(wdat[["vocab.teach.pre"]], wdat[["vocab.teach.pos"]])


ldat[["grupo:Cor.Raca"]] = wdat

(non.normal)
```

```{r}
aov = anova_test(wdat, vocab.teach.pos ~ vocab.teach.pre + grupo*Cor.Raca)
laov[["grupo:Cor.Raca"]] <- merge(get_anova_table(aov), laov[["grupo:Cor.Raca"]], by="Effect", suffixes = c("","'"))

(df = get_anova_table(aov))
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

```{r}
pwcs <- list()
pwcs[["Cor.Raca"]] <- emmeans_test(
  group_by(wdat, grupo), vocab.teach.pos ~ Cor.Raca,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")
pwcs[["grupo"]] <- emmeans_test(
  group_by(wdat, Cor.Raca), vocab.teach.pos ~ grupo,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")

pwc <- plyr::rbind.fill(pwcs[["grupo"]], pwcs[["Cor.Raca"]])
pwc <- pwc[,c("grupo","Cor.Raca", colnames(pwc)[!colnames(pwc) %in% c("grupo","Cor.Raca")])]
```

```{r, echo=FALSE, purl=FALSE}
kable(pwc, digits = 3)
```

```{r}
pwc.long <- emmeans_test(dplyr::group_by_at(wdat.long, c("grupo","Cor.Raca")),
                         vocab.teach ~ time,
                         p.adjust.method = "bonferroni")
lpwc[["grupo:Cor.Raca"]] <- merge(plyr::rbind.fill(pwc, pwc.long), lpwc[["grupo:Cor.Raca"]], by=c("grupo","Cor.Raca","term",".y.","group1","group2"), suffixes = c("","'"))
```

```{r, echo=FALSE, purl=FALSE}
kable(pwc.long, digits = 3)
```

```{r}
ds <- get.descriptives(wdat, "vocab.teach.pos", c("grupo","Cor.Raca"), covar = "vocab.teach.pre")
ds <- merge(ds[ds$variable != "vocab.teach.pre",],
            ds[ds$variable == "vocab.teach.pre", !colnames(ds) %in% c("variable")],
            by = c("grupo","Cor.Raca"), all.x = T, suffixes = c("", ".vocab.teach.pre"))
ds <- merge(get_emmeans(pwcs[["grupo"]]), ds, by = c("grupo","Cor.Raca"), suffixes = c(".emms", ""))
ds <- ds[,c("grupo","Cor.Raca","n","mean.vocab.teach.pre","se.vocab.teach.pre","mean","se","emmean","se.emms")]

colnames(ds) <- c("grupo","Cor.Raca", "N", paste0(c("M","SE")," (pre)"),
                  paste0(c("M","SE"), " (unadj)"), paste0(c("M", "SE"), " (adj)"))

lemms[["grupo:Cor.Raca"]] <- merge(ds, lemms[["grupo:Cor.Raca"]], by=c("grupo","Cor.Raca"), suffixes = c("","'"))
```

```{r, echo=FALSE, purl=FALSE}
kable(ds, digits = 3)
```


### Plots for ancova

```{r}
plots <- twoWayAncovaPlots(
  wdat, "vocab.teach.pos", c("grupo","Cor.Raca"), aov, pwcs, addParam = c("mean_se"),
  font.label.size=10, step.increase=0.05, p.label="p.adj",
  subtitle = which(aov$Effect == "grupo:Cor.Raca"))
```

```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(plots[["grupo"]]))
  plots[["grupo"]] + ggplot2::scale_color_manual(values = color[["Cor.Raca"]])
```


```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(plots[["Cor.Raca"]]))
  plots[["Cor.Raca"]] + ggplot2::scale_color_manual(values = color[["grupo"]])
```


```{r}
plots <- twoWayAncovaBoxPlots(
  wdat, "vocab.teach.pos", c("grupo","Cor.Raca"), aov, pwcs, covar = "vocab.teach.pre",
  theme = "classic", color = color[["grupo:Cor.Raca"]],
  subtitle = which(aov$Effect == "grupo:Cor.Raca"))
```

```{r, dpi=300, fig.width=8, fig.height=6}
plots[["grupo:Cor.Raca"]] + ggplot2::ylab("Vocabulario Ensinado") + ggplot2::scale_x_discrete(labels=c('pre', 'pos'))
```


```{r}
plots <- twoWayAncovaBoxPlots(
  wdat.long, "vocab.teach", c("grupo","Cor.Raca"), aov, pwc.long, pre.post = "time",
  theme = "classic", color = color$prepost)
```

```{r, dpi=300, fig.width=14, fig.height=6}
plots[["grupo:Cor.Raca"]] + ggplot2::ylab("Vocabulario Ensinado")
```


### Checking linearity assumption

```{r, dpi=300, fig.width=12, fig.height=8}
ggscatter(wdat, x = "vocab.teach.pre", y = "vocab.teach.pos", size = 0.5,
          facet.by = c("grupo","Cor.Raca"), add = "reg.line")+
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~"))
  )
```

### Checking normality and homogeneity

```{r}
res <- augment(lm(vocab.teach.pos ~ vocab.teach.pre + grupo*Cor.Raca, data = wdat))
```

```{r}
shapiro_test(res$.resid)
```

```{r}
levene_test(res, .resid ~ grupo*Cor.Raca)
```


# ANCOVA and Pairwise for two factors **grupo:Serie**

## Without remove non-normal data

```{r}
pdat = remove_group_data(dat[!is.na(dat[["grupo"]]) & !is.na(dat[["Serie"]]),], "vocab.teach.pos", c("grupo","Serie"))

pdat.long <- rbind(pdat[,c("id","grupo","Serie")], pdat[,c("id","grupo","Serie")])
pdat.long[["time"]] <- c(rep("pre", nrow(pdat)), rep("pos", nrow(pdat)))
pdat.long[["time"]] <- factor(pdat.long[["time"]], c("pre","pos"))
pdat.long[["vocab.teach"]] <- c(pdat[["vocab.teach.pre"]], pdat[["vocab.teach.pos"]])

aov = anova_test(pdat, vocab.teach.pos ~ vocab.teach.pre + grupo*Serie)
laov[["grupo:Serie"]] <- get_anova_table(aov)
```

```{r}
pwcs <- list()
pwcs[["Serie"]] <- emmeans_test(
  group_by(pdat, grupo), vocab.teach.pos ~ Serie,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")
pwcs[["grupo"]] <- emmeans_test(
  group_by(pdat, Serie), vocab.teach.pos ~ grupo,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")

pwc <- plyr::rbind.fill(pwcs[["grupo"]], pwcs[["Serie"]])
pwc <- pwc[,c("grupo","Serie", colnames(pwc)[!colnames(pwc) %in% c("grupo","Serie")])]
```

```{r}
pwc.long <- emmeans_test(dplyr::group_by_at(pdat.long, c("grupo","Serie")),
                         vocab.teach ~ time,
                         p.adjust.method = "bonferroni")
lpwc[["grupo:Serie"]] <- plyr::rbind.fill(pwc, pwc.long)
```


```{r}
ds <- get.descriptives(pdat, "vocab.teach.pos", c("grupo","Serie"), covar = "vocab.teach.pre")
ds <- merge(ds[ds$variable != "vocab.teach.pre",],
            ds[ds$variable == "vocab.teach.pre", !colnames(ds) %in% c("variable")],
            by = c("grupo","Serie"), all.x = T, suffixes = c("", ".vocab.teach.pre"))
ds <- merge(get_emmeans(pwcs[["grupo"]]), ds, by = c("grupo","Serie"), suffixes = c(".emms", ""))
ds <- ds[,c("grupo","Serie","n","mean.vocab.teach.pre","se.vocab.teach.pre","mean","se","emmean","se.emms")]

colnames(ds) <- c("grupo","Serie", "N", paste0(c("M","SE")," (pre)"),
                  paste0(c("M","SE"), " (unadj)"), paste0(c("M", "SE"), " (adj)"))

lemms[["grupo:Serie"]] <- ds
```

## Computing ANCOVA and PairWise After removing non-normal data (OK)

```{r}
wdat = pdat 

res = residuals(lm(vocab.teach.pos ~ vocab.teach.pre + grupo*Serie, data = wdat))
non.normal = getNonNormal(res, wdat$id, plimit = 0.05)

wdat = wdat[!wdat$id %in% non.normal,]

wdat.long <- rbind(wdat[,c("id","grupo","Serie")], wdat[,c("id","grupo","Serie")])
wdat.long[["time"]] <- c(rep("pre", nrow(wdat)), rep("pos", nrow(wdat)))
wdat.long[["time"]] <- factor(wdat.long[["time"]], c("pre","pos"))
wdat.long[["vocab.teach"]] <- c(wdat[["vocab.teach.pre"]], wdat[["vocab.teach.pos"]])


ldat[["grupo:Serie"]] = wdat

(non.normal)
```

```{r}
aov = anova_test(wdat, vocab.teach.pos ~ vocab.teach.pre + grupo*Serie)
laov[["grupo:Serie"]] <- merge(get_anova_table(aov), laov[["grupo:Serie"]], by="Effect", suffixes = c("","'"))

(df = get_anova_table(aov))
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

```{r}
pwcs <- list()
pwcs[["Serie"]] <- emmeans_test(
  group_by(wdat, grupo), vocab.teach.pos ~ Serie,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")
pwcs[["grupo"]] <- emmeans_test(
  group_by(wdat, Serie), vocab.teach.pos ~ grupo,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")

pwc <- plyr::rbind.fill(pwcs[["grupo"]], pwcs[["Serie"]])
pwc <- pwc[,c("grupo","Serie", colnames(pwc)[!colnames(pwc) %in% c("grupo","Serie")])]
```

```{r, echo=FALSE, purl=FALSE}
kable(pwc, digits = 3)
```

```{r}
pwc.long <- emmeans_test(dplyr::group_by_at(wdat.long, c("grupo","Serie")),
                         vocab.teach ~ time,
                         p.adjust.method = "bonferroni")
lpwc[["grupo:Serie"]] <- merge(plyr::rbind.fill(pwc, pwc.long), lpwc[["grupo:Serie"]], by=c("grupo","Serie","term",".y.","group1","group2"), suffixes = c("","'"))
```

```{r, echo=FALSE, purl=FALSE}
kable(pwc.long, digits = 3)
```

```{r}
ds <- get.descriptives(wdat, "vocab.teach.pos", c("grupo","Serie"), covar = "vocab.teach.pre")
ds <- merge(ds[ds$variable != "vocab.teach.pre",],
            ds[ds$variable == "vocab.teach.pre", !colnames(ds) %in% c("variable")],
            by = c("grupo","Serie"), all.x = T, suffixes = c("", ".vocab.teach.pre"))
ds <- merge(get_emmeans(pwcs[["grupo"]]), ds, by = c("grupo","Serie"), suffixes = c(".emms", ""))
ds <- ds[,c("grupo","Serie","n","mean.vocab.teach.pre","se.vocab.teach.pre","mean","se","emmean","se.emms")]

colnames(ds) <- c("grupo","Serie", "N", paste0(c("M","SE")," (pre)"),
                  paste0(c("M","SE"), " (unadj)"), paste0(c("M", "SE"), " (adj)"))

lemms[["grupo:Serie"]] <- merge(ds, lemms[["grupo:Serie"]], by=c("grupo","Serie"), suffixes = c("","'"))
```

```{r, echo=FALSE, purl=FALSE}
kable(ds, digits = 3)
```


### Plots for ancova

```{r}
plots <- twoWayAncovaPlots(
  wdat, "vocab.teach.pos", c("grupo","Serie"), aov, pwcs, addParam = c("mean_se"),
  font.label.size=10, step.increase=0.05, p.label="p.adj",
  subtitle = which(aov$Effect == "grupo:Serie"))
```

```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(plots[["grupo"]]))
  plots[["grupo"]] + ggplot2::scale_color_manual(values = color[["Serie"]])
```


```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(plots[["Serie"]]))
  plots[["Serie"]] + ggplot2::scale_color_manual(values = color[["grupo"]])
```


```{r}
plots <- twoWayAncovaBoxPlots(
  wdat, "vocab.teach.pos", c("grupo","Serie"), aov, pwcs, covar = "vocab.teach.pre",
  theme = "classic", color = color[["grupo:Serie"]],
  subtitle = which(aov$Effect == "grupo:Serie"))
```

```{r, dpi=300, fig.width=8, fig.height=6}
plots[["grupo:Serie"]] + ggplot2::ylab("Vocabulario Ensinado") + ggplot2::scale_x_discrete(labels=c('pre', 'pos'))
```


```{r}
plots <- twoWayAncovaBoxPlots(
  wdat.long, "vocab.teach", c("grupo","Serie"), aov, pwc.long, pre.post = "time",
  theme = "classic", color = color$prepost)
```

```{r, dpi=300, fig.width=14, fig.height=6}
plots[["grupo:Serie"]] + ggplot2::ylab("Vocabulario Ensinado")
```


### Checking linearity assumption

```{r, dpi=300, fig.width=12, fig.height=8}
ggscatter(wdat, x = "vocab.teach.pre", y = "vocab.teach.pos", size = 0.5,
          facet.by = c("grupo","Serie"), add = "reg.line")+
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~"))
  )
```

### Checking normality and homogeneity

```{r}
res <- augment(lm(vocab.teach.pos ~ vocab.teach.pre + grupo*Serie, data = wdat))
```

```{r}
shapiro_test(res$.resid)
```

```{r}
levene_test(res, .resid ~ grupo*Serie)
```


# ANCOVA and Pairwise for two factors **grupo:vocab.teach.quintile**

## Without remove non-normal data

```{r}
pdat = remove_group_data(dat[!is.na(dat[["grupo"]]) & !is.na(dat[["vocab.teach.quintile"]]),], "vocab.teach.pos", c("grupo","vocab.teach.quintile"))

pdat.long <- rbind(pdat[,c("id","grupo","vocab.teach.quintile")], pdat[,c("id","grupo","vocab.teach.quintile")])
pdat.long[["time"]] <- c(rep("pre", nrow(pdat)), rep("pos", nrow(pdat)))
pdat.long[["time"]] <- factor(pdat.long[["time"]], c("pre","pos"))
pdat.long[["vocab.teach"]] <- c(pdat[["vocab.teach.pre"]], pdat[["vocab.teach.pos"]])

aov = anova_test(pdat, vocab.teach.pos ~ vocab.teach.pre + grupo*vocab.teach.quintile)
laov[["grupo:vocab.teach.quintile"]] <- get_anova_table(aov)
```

```{r}
pwcs <- list()
pwcs[["vocab.teach.quintile"]] <- emmeans_test(
  group_by(pdat, grupo), vocab.teach.pos ~ vocab.teach.quintile,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")
pwcs[["grupo"]] <- emmeans_test(
  group_by(pdat, vocab.teach.quintile), vocab.teach.pos ~ grupo,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")

pwc <- plyr::rbind.fill(pwcs[["grupo"]], pwcs[["vocab.teach.quintile"]])
pwc <- pwc[,c("grupo","vocab.teach.quintile", colnames(pwc)[!colnames(pwc) %in% c("grupo","vocab.teach.quintile")])]
```

```{r}
pwc.long <- emmeans_test(dplyr::group_by_at(pdat.long, c("grupo","vocab.teach.quintile")),
                         vocab.teach ~ time,
                         p.adjust.method = "bonferroni")
lpwc[["grupo:vocab.teach.quintile"]] <- plyr::rbind.fill(pwc, pwc.long)
```


```{r}
ds <- get.descriptives(pdat, "vocab.teach.pos", c("grupo","vocab.teach.quintile"), covar = "vocab.teach.pre")
ds <- merge(ds[ds$variable != "vocab.teach.pre",],
            ds[ds$variable == "vocab.teach.pre", !colnames(ds) %in% c("variable")],
            by = c("grupo","vocab.teach.quintile"), all.x = T, suffixes = c("", ".vocab.teach.pre"))
ds <- merge(get_emmeans(pwcs[["grupo"]]), ds, by = c("grupo","vocab.teach.quintile"), suffixes = c(".emms", ""))
ds <- ds[,c("grupo","vocab.teach.quintile","n","mean.vocab.teach.pre","se.vocab.teach.pre","mean","se","emmean","se.emms")]

colnames(ds) <- c("grupo","vocab.teach.quintile", "N", paste0(c("M","SE")," (pre)"),
                  paste0(c("M","SE"), " (unadj)"), paste0(c("M", "SE"), " (adj)"))

lemms[["grupo:vocab.teach.quintile"]] <- ds
```

## Computing ANCOVA and PairWise After removing non-normal data (OK)

```{r}
wdat = pdat 

res = residuals(lm(vocab.teach.pos ~ vocab.teach.pre + grupo*vocab.teach.quintile, data = wdat))
non.normal = getNonNormal(res, wdat$id, plimit = 0.05)

wdat = wdat[!wdat$id %in% non.normal,]

wdat.long <- rbind(wdat[,c("id","grupo","vocab.teach.quintile")], wdat[,c("id","grupo","vocab.teach.quintile")])
wdat.long[["time"]] <- c(rep("pre", nrow(wdat)), rep("pos", nrow(wdat)))
wdat.long[["time"]] <- factor(wdat.long[["time"]], c("pre","pos"))
wdat.long[["vocab.teach"]] <- c(wdat[["vocab.teach.pre"]], wdat[["vocab.teach.pos"]])


ldat[["grupo:vocab.teach.quintile"]] = wdat

(non.normal)
```

```{r}
aov = anova_test(wdat, vocab.teach.pos ~ vocab.teach.pre + grupo*vocab.teach.quintile)
laov[["grupo:vocab.teach.quintile"]] <- merge(get_anova_table(aov), laov[["grupo:vocab.teach.quintile"]], by="Effect", suffixes = c("","'"))

(df = get_anova_table(aov))
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

```{r}
pwcs <- list()
pwcs[["vocab.teach.quintile"]] <- emmeans_test(
  group_by(wdat, grupo), vocab.teach.pos ~ vocab.teach.quintile,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")
pwcs[["grupo"]] <- emmeans_test(
  group_by(wdat, vocab.teach.quintile), vocab.teach.pos ~ grupo,
  covariate = vocab.teach.pre, p.adjust.method = "bonferroni")

pwc <- plyr::rbind.fill(pwcs[["grupo"]], pwcs[["vocab.teach.quintile"]])
pwc <- pwc[,c("grupo","vocab.teach.quintile", colnames(pwc)[!colnames(pwc) %in% c("grupo","vocab.teach.quintile")])]
```

```{r, echo=FALSE, purl=FALSE}
kable(pwc, digits = 3)
```

```{r}
pwc.long <- emmeans_test(dplyr::group_by_at(wdat.long, c("grupo","vocab.teach.quintile")),
                         vocab.teach ~ time,
                         p.adjust.method = "bonferroni")
lpwc[["grupo:vocab.teach.quintile"]] <- merge(plyr::rbind.fill(pwc, pwc.long), lpwc[["grupo:vocab.teach.quintile"]], by=c("grupo","vocab.teach.quintile","term",".y.","group1","group2"), suffixes = c("","'"))
```

```{r, echo=FALSE, purl=FALSE}
kable(pwc.long, digits = 3)
```

```{r}
ds <- get.descriptives(wdat, "vocab.teach.pos", c("grupo","vocab.teach.quintile"), covar = "vocab.teach.pre")
ds <- merge(ds[ds$variable != "vocab.teach.pre",],
            ds[ds$variable == "vocab.teach.pre", !colnames(ds) %in% c("variable")],
            by = c("grupo","vocab.teach.quintile"), all.x = T, suffixes = c("", ".vocab.teach.pre"))
ds <- merge(get_emmeans(pwcs[["grupo"]]), ds, by = c("grupo","vocab.teach.quintile"), suffixes = c(".emms", ""))
ds <- ds[,c("grupo","vocab.teach.quintile","n","mean.vocab.teach.pre","se.vocab.teach.pre","mean","se","emmean","se.emms")]

colnames(ds) <- c("grupo","vocab.teach.quintile", "N", paste0(c("M","SE")," (pre)"),
                  paste0(c("M","SE"), " (unadj)"), paste0(c("M", "SE"), " (adj)"))

lemms[["grupo:vocab.teach.quintile"]] <- merge(ds, lemms[["grupo:vocab.teach.quintile"]], by=c("grupo","vocab.teach.quintile"), suffixes = c("","'"))
```

```{r, echo=FALSE, purl=FALSE}
kable(ds, digits = 3)
```


### Plots for ancova

```{r}
plots <- twoWayAncovaPlots(
  wdat, "vocab.teach.pos", c("grupo","vocab.teach.quintile"), aov, pwcs, addParam = c("mean_se"),
  font.label.size=10, step.increase=0.05, p.label="p.adj",
  subtitle = which(aov$Effect == "grupo:vocab.teach.quintile"))
```

```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(plots[["grupo"]]))
  plots[["grupo"]] + ggplot2::scale_color_manual(values = color[["vocab.teach.quintile"]])
```


```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(plots[["vocab.teach.quintile"]]))
  plots[["vocab.teach.quintile"]] + ggplot2::scale_color_manual(values = color[["grupo"]])
```


```{r}
plots <- twoWayAncovaBoxPlots(
  wdat, "vocab.teach.pos", c("grupo","vocab.teach.quintile"), aov, pwcs, covar = "vocab.teach.pre",
  theme = "classic", color = color[["grupo:vocab.teach.quintile"]],
  subtitle = which(aov$Effect == "grupo:vocab.teach.quintile"))
```

```{r, dpi=300, fig.width=8, fig.height=6}
plots[["grupo:vocab.teach.quintile"]] + ggplot2::ylab("Vocabulario Ensinado") + ggplot2::scale_x_discrete(labels=c('pre', 'pos'))
```


```{r}
plots <- twoWayAncovaBoxPlots(
  wdat.long, "vocab.teach", c("grupo","vocab.teach.quintile"), aov, pwc.long, pre.post = "time",
  theme = "classic", color = color$prepost)
```

```{r, dpi=300, fig.width=14, fig.height=6}
plots[["grupo:vocab.teach.quintile"]] + ggplot2::ylab("Vocabulario Ensinado")
```


### Checking linearity assumption

```{r, dpi=300, fig.width=12, fig.height=8}
ggscatter(wdat, x = "vocab.teach.pre", y = "vocab.teach.pos", size = 0.5,
          facet.by = c("grupo","vocab.teach.quintile"), add = "reg.line")+
  stat_regline_equation(
    aes(label =  paste(..eq.label.., ..rr.label.., sep = "~~~~"))
  )
```

### Checking normality and homogeneity

```{r}
res <- augment(lm(vocab.teach.pos ~ vocab.teach.pre + grupo*vocab.teach.quintile, data = wdat))
```

```{r}
shapiro_test(res$.resid)
```

```{r}
levene_test(res, .resid ~ grupo*vocab.teach.quintile)
```



# Summary of Results

## Descriptive Statistics

```{r}
df <- get.descriptives(ldat[["grupo"]], c(dv.pre, dv.pos), c("grupo"), 
                       include.global = T, symmetry.test = T, normality.test = F)
df <- plyr::rbind.fill(
  df, do.call(plyr::rbind.fill, lapply(lfatores2, FUN = function(f) {
    if (nrow(dat) > 0 && sum(!is.na(unique(dat[[f]]))) > 1)
      get.descriptives(ldat[[paste0("grupo:",f)]], c(dv.pre,dv.pos), c("grupo", f),
                       symmetry.test = T, normality.test = F)
    }))
)
(df <- df[,c(fatores1[fatores1 %in% colnames(df)],"variable",
             colnames(df)[!colnames(df) %in% c(fatores1,"variable")])])
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```


## ANCOVA Table Comparison

```{r}
df <- do.call(plyr::rbind.fill, laov)
(df <- df[!duplicated(df$Effect),])
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```


## PairWise Table Comparison

```{r}
df <- do.call(plyr::rbind.fill, lpwc)
df <- df[,c(names(lfatores), names(df)[!names(df) %in% c(names(lfatores),"term",".y.")])]
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```


## EMMS Table Comparison

```{r}
df <- do.call(plyr::rbind.fill, lemms)
df[["N-N'"]] <- df[["N"]] - df[["N'"]]
df <- df[,c(names(lfatores), names(df)[!names(df) %in% names(lfatores)])]
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```


