---
title: "ANOVA in Gains for gain Vocabulario (acertos) (gain Vocabulario (acertos))"
author: Geiser C. Challco <geiser@alumni.usp.br>
comment: Teste ANCOVA para determinar se houve diferenças significativas
         na gain Vocabulario (acertos) (medido usando pre- e pos-testes).
         
         ANCOVA test to determine whether there were significant differences
         in gain Vocabulario (acertos) (measured using pre- and post-tests).
         
         Author - Geiser C. Challco <geiser@alumni.usp.br>
         
         Shiny-Statistic is distributed in the hope that it will be useful,
         but WITHOUT ANY WARRANTY; without even the implied warranty of
         MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
         GNU General Public License for more details.
         
         You should have received a copy of the GNU General Public License.
         This file is  generate using Shiny-Statistic app (https://github.com/geiser/rshinystatistics)
         If not, see <https://www.gnu.org/licenses/>.
output:
  github_document:
    toc: true
  word_document:
    toc: true
  html_document:
    toc: true
fontsize: 10pt
---


**NOTE**

 - Teste ANOVA para determinar se houve diferenças significativas
no gain Vocabulario (acertos) (medido usando a diferença entre post-test e pre-testes).
 - ANOVA test to determine whether there were significant differences
in gain Vocabulario (acertos) (measured using the difference between post-test and pre-tests).

```{r setup, include=FALSE}
# Install and Load Packages
if (!'remotes' %in% rownames(installed.packages())) install.packages('remotes')
if (!"rshinystatistics" %in% rownames(installed.packages())) {
  remotes::install_github("geiser/rshinystatistics")
} else if (packageVersion("rshinystatistics") < "0.0.0.9900") {
  remotes::install_github("geiser/rshinystatistics")
}

wants <- c('ggplot2','ggpubr','rshinystatistics','utils','randomcoloR')
has <- wants %in% rownames(installed.packages())
if (any(!has)) install.packages(wants[!has])

library(readxl)

library(shiny)
library(esquisse)
library(scales)
library(knitr)
library(rmarkdown)

library(utils)
library(ggpubr)
library(ggplot2)
library(randomcoloR)

library(dplyr)

library(rshinystatistics)
```

```{r, include=FALSE, purl=FALSE}
options(knitr.kable.NA = '')
opts_chunk$set(echo = FALSE)
defaultW <- getOption("warn")
options(warn = -1)
```

```{r, include=FALSE}
# Setting Initial Variables
dv = "gain.vocab"

fatores2 <- c("Sexo","Zona","Cor.Raca")
lfatores2 <- as.list(fatores2)
names(lfatores2) <- fatores2

fatores1 <- c("grupo", fatores2)
lfatores1 <- as.list(fatores1)
names(lfatores1) <- fatores1

lfatores <- c(lfatores1)

# ..

gdat <- read_excel("../data/data.xlsx", sheet = "sumary")
gdat <- gdat[which(is.na(gdat$Necessidade.Deficiencia) & !is.na(gdat$Stari.Grupo)),]



dat <- gdat
dat$grupo <- factor(dat[["Stari.Grupo"]], c("Controle","Experimental"))
for (coln in c(names(lfatores))) {
  dat[[coln]] <- factor(dat[[coln]], sort(unique(dat[[coln]])[!is.na(unique(dat[[coln]]))]))
}
dat <- dat[which(!is.na(dat[[dv]])),]
dat <- dat[,c("id",names(lfatores),dv)]

color <- list()
color[["grupo"]] = c("#bcbd22","#fd7f6f")
color[["Sexo"]] = c("#FF007F","#4D4DFF")
color[["Zona"]] = c("#AA00FF","#00CCCC")
color[["Cor.Raca"]] = c(
  "Parda"="#b97100","Indígena"="#9F262F",
  "Branca"="#87c498", "Preta"="#848283","Amarela"="#D6B91C"
)

color[["grupo:Sexo"]] = c(
  "Controle:F"="#ff99cb", "Controle:M"="#b7b7ff",
  "Experimental:F"="#FF007F", "Experimental:M"="#4D4DFF",
  "Controle.F"="#ff99cb", "Controle.M"="#b7b7ff",
  "Experimental.F"="#FF007F", "Experimental.M"="#4D4DFF"
)
color[["grupo:Zona"]] = c(
  "Controle:Rural"="#b2efef","Controle:Urbana"="#e5b2ff",
  "Experimental:Rural"="#00CCCC", "Experimental:Urbana"="#AA00FF",
  "Controle.Rural"="#b2efef","Controle.Urbana"="#e5b2ff",
  "Experimental.Rural"="#00CCCC", "Experimental.Urbana"="#AA00FF"
)
color[["grupo:Cor.Raca"]] = c(
    "Controle:Parda"="#e3c699", "Experimental:Parda"="#b97100",
    "Controle:Indígena"="#e2bdc0", "Experimental:Indígena"="#9F262F",
    "Controle:Branca"="#c0e8cb", "Experimental:Branca"="#87c498",
    "Controle:Preta"="#dad9d9", "Experimental:Preta"="#848283",
    "Controle:Amarela"="#eee3a4", "Experimental:Amarela"="#D6B91C",
    
    "Controle.Parda"="#e3c699", "Experimental.Parda"="#b97100",
    "Controle.Indígena"="#e2bdc0", "Experimental.Indígena"="#9F262F",
    "Controle.Branca"="#c0e8cb", "Experimental.Branca"="#87c498",
    "Controle.Preta"="#dad9d9", "Experimental.Preta"="#848283",
    "Controle.Amarela"="#eee3a4", "Experimental.Amarela"="#D6B91C"
)


for (coln in c("vocab","score.tde",
               "TFL.lidas.per.min","TFL.corretas.per.min","leitura.compreensao")) {
  color[[paste0(coln,".quintile")]] = c("#BF0040","#FF0000","#800080","#0000FF","#4000BF")
  color[[paste0("grupo:",coln,".quintile")]] = c(
    "Experimental.1st quintile"="#BF0040", "Controle.1st quintile"="#d8668c",
    "Experimental.2nd quintile"="#FF0000", "Controle.2nd quintile"="#ff7f7f",
    "Experimental.3rd quintile"="#8fce00", "Controle.3rd quintile"="#ddf0b2",
    "Experimental.4th quintile"="#0000FF", "Controle.4th quintile"="#b2b2ff",
    "Experimental.5th quintile"="#4000BF", "Controle.5th quintile"="#b299e5",
    
    "Experimental:1st quintile"="#BF0040", "Controle:1st quintile"="#d8668c",
    "Experimental:2nd quintile"="#FF0000", "Controle:2nd quintile"="#ff7f7f",
    "Experimental:3rd quintile"="#8fce00", "Controle:3rd quintile"="#ddf0b2",
    "Experimental:4th quintile"="#0000FF", "Controle:4th quintile"="#b2b2ff",
    "Experimental:5th quintile"="#4000BF", "Controle:5th quintile"="#b299e5")
}


for (f in c("grupo", names(lfatores))) {
  if (is.null(color[[f]]) && length(unique(dat[[f]])) > 0) 
      color[[f]] <- distinctColorPalette(length(unique(dat[[f]])))
}
for (f in c(fatores2)) {
  if (is.null(color[[paste0("grupo:",f)]]) && length(unique(dat[[f]])) > 0)
    color[[paste0("grupo:",f)]] <- distinctColorPalette(length(unique(dat[["grupo"]]))*length(unique(dat[[f]])))
}
```

# Descriptive Statistics of Initial Data

```{r, include=FALSE}
df <- do.call(plyr::rbind.fill, lapply(lfatores2, FUN = function(f) {
  if (nrow(dat) > 0 && sum(!is.na(unique(dat[[f]]))) > 1)
    get.descriptives(dat, dv, c("grupo", f))
}))
(df <- df[,c(fatores1[fatores1 %in% colnames(df)],"variable",
             colnames(df)[!colnames(df) %in% c(fatores1,"variable")])])
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

# Checking of Assumptions

## Assumption: Normality distribution of data

```{r, echo=FALSE, include=FALSE}
(df <- do.call(plyr::rbind.fill, lapply(lfatores2, FUN = function(f) {
  if (nrow(dat) > 0 && sum(!is.na(unique(dat[[f]]))) > 1)
    normality.test.by.residual(dat,dv,c("grupo", f))
})))
```

```{r, echo=FALSE, purl=FALSE}
kable(df)
```

## Assumption: Homogeneity of data distribution

```{r, echo=FALSE, include=FALSE}
(df <- do.call(plyr::rbind.fill, lapply(lfatores2, FUN = function(f) {
  if (nrow(dat) > 0 && sum(!is.na(unique(dat[[f]]))) > 1)
    homogeneity.test(dat[which(!is.na(dat[[f]])),],dv,c("grupo",f))
})))
```

```{r, echo=FALSE, purl=FALSE}
kable(df)
```

# Computation of ANCOVA test and Pairwise Comparison

## ANCOVA tests for one factor

```{r, echo=FALSE, include=FALSE}
aovs1 <- lapply(lfatores1, FUN = function(x) {
  aov = anova.test(dat[!is.na(dat[[x]]),],dv,x,type=2)[[dv]]
  if (!is.null(aov))
    cbind(formula=rep(paste0("gain.vocab~",x),nrow(aov)),aov)
})

aov1 <- do.call(rbind, aovs1)
row.names(aov1) <- c()
(aov1 <- aov1[!duplicated(aov1$Effect),])
```

```{r, echo=FALSE, purl=FALSE}
df <- round.pval(aov1)
kable(df[,c("Effect","DFn","DFd","SSn","SSd","F","p","ges","p<.05")], digits=3)
```

## ANCOVA tests for two factors

```{r, echo=FALSE, include=FALSE}
aovs2 <- lapply(lfatores2, FUN = function(x) {
  tdat <- dat[!is.na(dat[[x]]),]
  if (length(unique(tdat$grupo)) >= 2 &&  length(unique(tdat[[x]])) >= 2) {
    aov = anova.test(tdat,dv,c("grupo",x),type=2)[[dv]]
    if (!is.null(aov))
      return(cbind(formula=rep(paste0("gain.vocab~grupo","*",x),nrow(aov)), aov))  
  }
})

aov2 <- do.call(rbind, aovs2)
row.names(aov2) <- c()
(aov2 <- aov2[!duplicated(aov2$Effect) & !aov2$Effect %in% fatores1,])
```

```{r, echo=FALSE, purl=FALSE}
df <- round.pval(aov2)
kable(df[,c("Effect","DFn","DFd","SSn","SSd","F","p","ges","p<.05")], digits=3)
```



## Pairwise comparisons for one factor: **grupo** 

```{r, include=FALSE}
pdat = dat[!is.na(dat[["grupo"]]),]
pdat = tryCatch(remove_group_data(pdat, "gain.vocab", c("grupo")), error = function(e) NULL)

if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1) {
  pwc = anova.pwc(pdat, "gain.vocab", "grupo")
  (apa <- get.anova.emmeans.with.ds(pwc, pdat, "gain.vocab", c("grupo"), "apa-format"))
}
```

```{r, echo=FALSE, purl=FALSE}
if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1)
  kable(apa, digits = 3)
```

```{r, include=FALSE}
if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1) {
  pdf <- get.anova.pwc.table(pwc, only.sig = F)
  (pdf <- pdf[!is.na(pdf$p.adj),])
}
```

```{r, echo=FALSE, purl=FALSE}
if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1) {
  pdf <- round.pval(pdf)
  kable(pdf[,c(".y.","group1","group2","estimate","conf.low",
               "conf.high","se","statistic","p","p.adj","p.adj.signif")], digits = 3)
}
```


```{r}
if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1)
  plots <- oneWayAnovaPlots(pdat, "gain.vocab", "grupo", aov1, pwc$gain.vocab,
                            addParam = c("mean_se"), font.label.size=10,
                            step.increase=0.05, p.label="p.adj",
                            subtitle = which(aov1$Effect == "grupo"))
```

```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(pdat) && !is.null(nrow(plots[["grupo"]]$data)))
  plots[["grupo"]] + ggplot2::scale_color_manual(values = color[["grupo"]])
```



```{r}
#if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1)
#  plots <- oneWayAncovaBoxPlots(
#    pdat, "gain.vocab", "grupo", aov1, pwc$gain.vocab,",
#    theme = "classic", color = color[["grupo"]],
#    subtitle = which(aov1$Effect == "grupo"))
```

```{r, dpi=300, fig.width=8, fig.height=6}
#if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1)
#  plots[["grupo"]] + ggplot2::ylab("gain Vocabulario (acertos)") + ggplot2::scale_x_discrete(labels=c('pre', 'pos'))
```


```{r}
#if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1)
#  plots <- oneWayAncovaBoxPlots(
#    pdat.long, "gain.vocab", "grupo", aov1, pwc.long, pre.post = "time",
#    theme = "classic", color = color$prepost)
```

```{r, dpi=300, fig.width=10, fig.height=6}
#if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1)
#  plots[["grupo"]] + ggplot2::ylab("gain Vocabulario (acertos)")
```



## Pairwise comparisons for two factors



### factores: **grupo:Sexo**

```{r, include=FALSE}
pdat = dat[!is.na(dat[["grupo"]]) & !is.na(dat[["Sexo"]]),]
pdat = tryCatch(remove_group_data(pdat, "gain.vocab", c("grupo","Sexo")), error = function(e) NULL)

if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Sexo"]])) > 1) {
  pwc = anova.pwc(pdat, "gain.vocab", c("grupo","Sexo"))
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Sexo"]]))
    (apa <- get.anova.emmeans.with.ds(pwc, pdat, "gain.vocab", c("grupo","Sexo"), "apa-format"))
}
```

```{r, echo=FALSE, purl=FALSE}
if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Sexo"]])) > 1)
  kable(apa, digits = 3)
```

```{r, include=FALSE}
if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Sexo"]])) > 1) {
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Sexo"]])) {
    pdf <- get.anova.pwc.table(pwc, only.sig = F)
    (pdf <- pdf[!is.na(pdf$p.adj),])
  } else 
    pdf <- NULL
}
```

```{r, echo=FALSE, purl=FALSE}
if (!is.null(pdf) && !is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Sexo"]])) > 1) {
  pdf <- round.pval(pdf)
  kable(pdf[,c(".y.","grupo","Sexo","group1","group2","estimate","conf.low",
               "conf.high","se","statistic","p","p.adj","p.adj.signif")], digits = 3)
}
```


```{r}
if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Sexo"]])) > 1)
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Sexo"]]))
    plots <- twoWayAnovaPlots(
      pdat, "gain.vocab", c("grupo","Sexo"), aov2, pwc$gain.vocab,
      addParam = c("mean_se"), font.label.size=10,
      step.increase=0.05, subtitle = which(aov2$Effect == "grupo:Sexo"))
```

```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(pdat) && !is.null(nrow(plots[["grupo"]]$data)))
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Sexo"]]))
    plots[["grupo"]] + ggplot2::scale_color_manual(values = color[["Sexo"]])
```

```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(pdat) && !is.null(nrow(plots[["Sexo"]]$data)))
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Sexo"]]))
    plots[["Sexo"]] + ggplot2::scale_color_manual(values = color[["grupo"]])
```


```{r}
#if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Sexo"]])) > 1)
#  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Sexo"]]))
#    plots <- twoWayAncovaBoxPlots(
#      pdat, "gain.vocab", c("grupo","Sexo"), aov2, pwc$gain.vocab,
#      theme = "classic", color = color[["grupo:Sexo"]],
#      subtitle = which(aov2$Effect == "grupo:Sexo"))
```

```{r, dpi=300, fig.width=8, fig.height=6}
#if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Sexo"]])) > 1)
#  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Sexo"]]))
#    plots[["grupo:Sexo"]] + ggplot2::ylab("gain Vocabulario (acertos)") + ggplot2::scale_x_discrete(labels=c('pre', 'pos'))
```

### factores: **grupo:Zona**

```{r, include=FALSE}
pdat = dat[!is.na(dat[["grupo"]]) & !is.na(dat[["Zona"]]),]
pdat = tryCatch(remove_group_data(pdat, "gain.vocab", c("grupo","Zona")), error = function(e) NULL)

if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Zona"]])) > 1) {
  pwc = anova.pwc(pdat, "gain.vocab", c("grupo","Zona"))
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Zona"]]))
    (apa <- get.anova.emmeans.with.ds(pwc, pdat, "gain.vocab", c("grupo","Zona"), "apa-format"))
}
```

```{r, echo=FALSE, purl=FALSE}
if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Zona"]])) > 1)
  kable(apa, digits = 3)
```

```{r, include=FALSE}
if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Zona"]])) > 1) {
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Zona"]])) {
    pdf <- get.anova.pwc.table(pwc, only.sig = F)
    (pdf <- pdf[!is.na(pdf$p.adj),])
  } else 
    pdf <- NULL
}
```

```{r, echo=FALSE, purl=FALSE}
if (!is.null(pdf) && !is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Zona"]])) > 1) {
  pdf <- round.pval(pdf)
  kable(pdf[,c(".y.","grupo","Zona","group1","group2","estimate","conf.low",
               "conf.high","se","statistic","p","p.adj","p.adj.signif")], digits = 3)
}
```


```{r}
if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Zona"]])) > 1)
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Zona"]]))
    plots <- twoWayAnovaPlots(
      pdat, "gain.vocab", c("grupo","Zona"), aov2, pwc$gain.vocab,
      addParam = c("mean_se"), font.label.size=10,
      step.increase=0.05, subtitle = which(aov2$Effect == "grupo:Zona"))
```

```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(pdat) && !is.null(nrow(plots[["grupo"]]$data)))
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Zona"]]))
    plots[["grupo"]] + ggplot2::scale_color_manual(values = color[["Zona"]])
```

```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(pdat) && !is.null(nrow(plots[["Zona"]]$data)))
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Zona"]]))
    plots[["Zona"]] + ggplot2::scale_color_manual(values = color[["grupo"]])
```


```{r}
#if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Zona"]])) > 1)
#  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Zona"]]))
#    plots <- twoWayAncovaBoxPlots(
#      pdat, "gain.vocab", c("grupo","Zona"), aov2, pwc$gain.vocab,
#      theme = "classic", color = color[["grupo:Zona"]],
#      subtitle = which(aov2$Effect == "grupo:Zona"))
```

```{r, dpi=300, fig.width=8, fig.height=6}
#if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Zona"]])) > 1)
#  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Zona"]]))
#    plots[["grupo:Zona"]] + ggplot2::ylab("gain Vocabulario (acertos)") + ggplot2::scale_x_discrete(labels=c('pre', 'pos'))
```

### factores: **grupo:Cor.Raca**

```{r, include=FALSE}
pdat = dat[!is.na(dat[["grupo"]]) & !is.na(dat[["Cor.Raca"]]),]
pdat = tryCatch(remove_group_data(pdat, "gain.vocab", c("grupo","Cor.Raca")), error = function(e) NULL)

if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Cor.Raca"]])) > 1) {
  pwc = anova.pwc(pdat, "gain.vocab", c("grupo","Cor.Raca"))
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Cor.Raca"]]))
    (apa <- get.anova.emmeans.with.ds(pwc, pdat, "gain.vocab", c("grupo","Cor.Raca"), "apa-format"))
}
```

```{r, echo=FALSE, purl=FALSE}
if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Cor.Raca"]])) > 1)
  kable(apa, digits = 3)
```

```{r, include=FALSE}
if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Cor.Raca"]])) > 1) {
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Cor.Raca"]])) {
    pdf <- get.anova.pwc.table(pwc, only.sig = F)
    (pdf <- pdf[!is.na(pdf$p.adj),])
  } else 
    pdf <- NULL
}
```

```{r, echo=FALSE, purl=FALSE}
if (!is.null(pdf) && !is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Cor.Raca"]])) > 1) {
  pdf <- round.pval(pdf)
  kable(pdf[,c(".y.","grupo","Cor.Raca","group1","group2","estimate","conf.low",
               "conf.high","se","statistic","p","p.adj","p.adj.signif")], digits = 3)
}
```


```{r}
if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Cor.Raca"]])) > 1)
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Cor.Raca"]]))
    plots <- twoWayAnovaPlots(
      pdat, "gain.vocab", c("grupo","Cor.Raca"), aov2, pwc$gain.vocab,
      addParam = c("mean_se"), font.label.size=10,
      step.increase=0.05, subtitle = which(aov2$Effect == "grupo:Cor.Raca"))
```

```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(pdat) && !is.null(nrow(plots[["grupo"]]$data)))
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Cor.Raca"]]))
    plots[["grupo"]] + ggplot2::scale_color_manual(values = color[["Cor.Raca"]])
```

```{r, dpi=300, fig.width=7, fig.height=7}
if (!is.null(pdat) && !is.null(nrow(plots[["Cor.Raca"]]$data)))
  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Cor.Raca"]]))
    plots[["Cor.Raca"]] + ggplot2::scale_color_manual(values = color[["grupo"]])
```


```{r}
#if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Cor.Raca"]])) > 1)
#  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Cor.Raca"]]))
#    plots <- twoWayAncovaBoxPlots(
#      pdat, "gain.vocab", c("grupo","Cor.Raca"), aov2, pwc$gain.vocab,
#      theme = "classic", color = color[["grupo:Cor.Raca"]],
#      subtitle = which(aov2$Effect == "grupo:Cor.Raca"))
```

```{r, dpi=300, fig.width=8, fig.height=6}
#if (!is.null(pdat) && length(unique(pdat[["grupo"]])) > 1 && length(unique(pdat[["Cor.Raca"]])) > 1)
#  if (!is.null(pwc[["gain.vocab"]][["grupo"]]) || !is.null(pwc[["gain.vocab"]][["Cor.Raca"]]))
#    plots[["grupo:Cor.Raca"]] + ggplot2::ylab("gain Vocabulario (acertos)") + ggplot2::scale_x_discrete(labels=c('pre', 'pos'))
```
