---
title: "ANCOVA in {{ label.en }} ({{ label.pt }})"
author: Geiser C. Challco <geiser@alumni.usp.br>
comment: Teste ANCOVA para determinar se houve diferenças significativas
         na {{ label.pt }} (medido usando pre- e pos-testes).
         
         ANCOVA test to determine whether there were significant differences
         in {{ label.en }} (measured using pre- and post-tests).
         
         Author - Geiser C. Challco <geiser@alumni.usp.br>
         
         Shiny-Statistic is distributed in the hope that it will be useful,
         but WITHOUT ANY WARRANTY; without even the implied warranty of
         MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
         GNU General Public License for more details.
         
         You should have received a copy of the GNU General Public License.
         This file is  generate using Shiny-Statistic app (https://github.com/geiser/rshinystatistics)
         If not, see <https://www.gnu.org/licenses/>.
output:
  github_document:
    toc: true
  word_document:
    toc: true
  html_document:
    toc: true
fontsize: 10pt
---


**NOTE**

 - Teste ANCOVA para determinar se houve diferenças significativas
no {{ label.pt }} (medido usando pre- e pos-testes).
 - ANCOVA test to determine whether there were significant differences
in {{ label.en }} (measured using pre- and post-tests).

```{r setup, include=FALSE}
# Install and Load Packages
if (!'remotes' %in% rownames(installed.packages())) install.packages('remotes')
if (!"rshinystatistics" %in% rownames(installed.packages())) {
  remotes::install_github("geiser/rshinystatistics")
} else if (packageVersion("rshinystatistics") < "0.0.0.9900") {
  remotes::install_github("geiser/rshinystatistics")
}

wants <- c('ggplot2','ggpubr','rshinystatistics','utils','randomcoloR')
has <- wants %in% rownames(installed.packages())
if (any(!has)) install.packages(wants[!has])

library(readxl)

library(shiny)
library(esquisse)
library(scales)
library(knitr)
library(rmarkdown)

library(utils)
library(ggpubr)
library(ggplot2)
library(randomcoloR)

library(dplyr)

library(rshinystatistics)
```

```{r, include=FALSE, purl=FALSE}
options(knitr.kable.NA = '')
opts_chunk$set(echo = FALSE)
defaultW <- getOption("warn")
options(warn = -1)
```

```{r, include=FALSE}
# Setting Initial Variables
dv = "{{ dv }}"
dv.pos = "{{ dv.pos }}"
dv.pre = "{{ dv.pre }}"

fatores2 <- c({{ paste0(paste0('"', fatores ,'"'), collapse = ",") }})
lfatores2 <- as.list(fatores2)
names(lfatores2) <- fatores2

fatores1 <- c("grupo", fatores2)
lfatores1 <- as.list(fatores1)
names(lfatores1) <- fatores1

lfatores <- c(lfatores1)

# ..

gdat <- read_excel("data/data.xlsx", sheet = "sumary")
gdat <- gdat[which(is.na(gdat$Necessidade.Deficiencia) & !is.na(gdat${{ grupo }})),]

{{ dat.filter }}

dat <- gdat
dat$grupo <- factor(dat[["{{ grupo }}"]], c("Controle","Experimental"))
for (coln in c(names(lfatores))) {
  dat[[coln]] <- factor(dat[[coln]], sort(unique(dat[[coln]])[!is.na(unique(dat[[coln]]))]))
}
dat <- dat[which(!is.na(dat[[dv.pre]]) & !is.na(dat[[dv.pos]])),]
dat <- dat[,c("id",names(lfatores),dv.pre,dv.pos)]

dat.long <- rbind(dat, dat)
dat.long$time <- c(rep("pre", nrow(dat)), rep("pos", nrow(dat)))
dat.long$time <- factor(dat.long$time, c("pre","pos"))
dat.long[[dv]] <- c(dat[[dv.pre]], dat[[dv.pos]])

color <- list()
color[["prepost"]] = c("#ffee65","#f28e2B")
color[["grupo"]] = c("#bcbd22","{{ exp.color }}")
color[["Sexo"]] = c("#FF007F","#4D4DFF")
color[["Zona"]] = c("#AA00FF","#00CCCC")
color[["Cor.Raca"]] = c(
  "Parda"="#b97100","Indígena"="#9F262F",
  "Branca"="#87c498", "Preta"="#848283","Amarela"="#D6B91C"
)

color[["grupo:Sexo"]] = c(
  "Controle:F"="#ff99cb", "Controle:M"="#b7b7ff",
  "Experimental:F"="#FF007F", "Experimental:M"="#4D4DFF",
  "Controle.F"="#ff99cb", "Controle.M"="#b7b7ff",
  "Experimental.F"="#FF007F", "Experimental.M"="#4D4DFF"
)
color[["grupo:Zona"]] = c(
  "Controle:Rural"="#b2efef","Controle:Urbana"="#e5b2ff",
  "Experimental:Rural"="#00CCCC", "Experimental:Urbana"="#AA00FF",
  "Controle.Rural"="#b2efef","Controle.Urbana"="#e5b2ff",
  "Experimental.Rural"="#00CCCC", "Experimental.Urbana"="#AA00FF"
)
color[["grupo:Cor.Raca"]] = c(
    "Controle:Parda"="#e3c699", "Experimental:Parda"="#b97100",
    "Controle:Indígena"="#e2bdc0", "Experimental:Indígena"="#9F262F",
    "Controle:Branca"="#c0e8cb", "Experimental:Branca"="#87c498",
    "Controle:Preta"="#dad9d9", "Experimental:Preta"="#848283",
    "Controle:Amarela"="#eee3a4", "Experimental:Amarela"="#D6B91C",
    
    "Controle.Parda"="#e3c699", "Experimental.Parda"="#b97100",
    "Controle.Indígena"="#e2bdc0", "Experimental.Indígena"="#9F262F",
    "Controle.Branca"="#c0e8cb", "Experimental.Branca"="#87c498",
    "Controle.Preta"="#dad9d9", "Experimental.Preta"="#848283",
    "Controle.Amarela"="#eee3a4", "Experimental.Amarela"="#D6B91C"
)


for (coln in c("vocab","score.tde",
               "TFL.lidas.per.min","TFL.corretas.per.min","leitura.compreensao")) {
  color[[paste0(coln,".quintile")]] = c("#BF0040","#FF0000","#800080","#0000FF","#4000BF")
  color[[paste0("grupo:",coln,".quintile")]] = c(
    "Experimental.1st quintile"="#BF0040", "Controle.1st quintile"="#d8668c",
    "Experimental.2nd quintile"="#FF0000", "Controle.2nd quintile"="#ff7f7f",
    "Experimental.3rd quintile"="#8fce00", "Controle.3rd quintile"="#ddf0b2",
    "Experimental.4th quintile"="#0000FF", "Controle.4th quintile"="#b2b2ff",
    "Experimental.5th quintile"="#4000BF", "Controle.5th quintile"="#b299e5",
    
    "Experimental:1st quintile"="#BF0040", "Controle:1st quintile"="#d8668c",
    "Experimental:2nd quintile"="#FF0000", "Controle:2nd quintile"="#ff7f7f",
    "Experimental:3rd quintile"="#8fce00", "Controle:3rd quintile"="#ddf0b2",
    "Experimental:4th quintile"="#0000FF", "Controle:4th quintile"="#b2b2ff",
    "Experimental:5th quintile"="#4000BF", "Controle:5th quintile"="#b299e5")
}


for (f in c("grupo", names(lfatores))) {
  if (is.null(color[[f]]) && length(unique(dat[[f]])) > 0) 
      color[[f]] <- distinctColorPalette(length(unique(dat[[f]])))
}
for (f in c(fatores2)) {
  if (is.null(color[[paste0("grupo:",f)]]) && length(unique(dat[[f]])) > 0)
    color[[paste0("grupo:",f)]] <- distinctColorPalette(length(unique(dat[["grupo"]]))*length(unique(dat[[f]])))
}
```

# Descriptive Statistics of Initial Data

```{r, include=FALSE}
df <- do.call(plyr::rbind.fill, lapply(lfatores2, FUN = function(f) {
  if (nrow(dat) > 0 && sum(!is.na(unique(dat[[f]]))) > 1)
    get.descriptives(dat, c(dv.pre,dv.pos), c("grupo", f))
}))
(df <- df[,c(fatores1[fatores1 %in% colnames(df)],"variable",
             colnames(df)[!colnames(df) %in% c(fatores1,"variable")])])
```

```{r, echo=FALSE, purl=FALSE}
kable(df, digits = 3)
```

# Checking of Assumptions

## Assumption: Normality distribution of data

```{r, echo=FALSE, include=FALSE}
(df <- do.call(plyr::rbind.fill, lapply(lfatores2, FUN = function(f) {
  if (nrow(dat) > 0 && sum(!is.na(unique(dat[[f]]))) > 1)
    normality.test.by.residual(dat,dv.pos,c("grupo", f),c(),dv.pre)
})))
```

```{r, echo=FALSE, purl=FALSE}
kable(df)
```

## Assumption: Homogeneity of data distribution

```{r, echo=FALSE, include=FALSE}
(df <- do.call(plyr::rbind.fill, lapply(lfatores2, FUN = function(f) {
  if (nrow(dat) > 0 && sum(!is.na(unique(dat[[f]]))) > 1)
    homogeneity.test(dat[which(!is.na(dat[[f]])),],dv.pos,c("grupo",f),c(),dv.pre)
})))
```

```{r, echo=FALSE, purl=FALSE}
kable(df)
```

# Computation of ANCOVA test and Pairwise Comparison

## ANCOVA tests for one factor

```{r, echo=FALSE, include=FALSE}
aovs1 <- lapply(lfatores1, FUN = function(x) {
  aov = ancova.test(dat[!is.na(dat[[x]]),],dv.pos,x,dv.pre,2,"ges")[[dv.pos]]
  if (!is.null(aov))
    cbind(formula=rep(paste0("{{ dv.pos }}~{{ dv.pre }}+",x),nrow(aov)),aov)
})

aov1 <- do.call(rbind, aovs1)
row.names(aov1) <- c()
(aov1 <- aov1[!duplicated(aov1$Effect),])
```

```{r, echo=FALSE, purl=FALSE}
df <- round.pval(aov1)
kable(df[,c("Effect","DFn","DFd","SSn","SSd","F","p","ges","p<.05")], digits=3)
```

## ANCOVA tests for two factors

```{r, echo=FALSE, include=FALSE}
aovs2 <- lapply(lfatores2, FUN = function(x) {
  tdat <- dat[!is.na(dat[[x]]),]
  if (length(unique(tdat$grupo)) >= 2 &&  length(unique(tdat[[x]])) >= 2) {
    aov = ancova.test(tdat,dv.pos,c("grupo",x),dv.pre, 2,"ges")[[dv.pos]]
    if (!is.null(aov))
      return(cbind(formula=rep(paste0("{{ dv.pos }}~{{ dv.pre }}+grupo","*",x),nrow(aov)), aov))  
  }
})

aov2 <- do.call(rbind, aovs2)
row.names(aov2) <- c()
(aov2 <- aov2[!duplicated(aov2$Effect) & !aov2$Effect %in% fatores1,])
```

```{r, echo=FALSE, purl=FALSE}
df <- round.pval(aov2)
kable(df[,c("Effect","DFn","DFd","SSn","SSd","F","p","ges","p<.05")], digits=3)
```


